The Development of a Collaborative Tool to Teach Debugging

### 2020-12-02 paper/paper.tex
Touch of program cleanup discovered when writing the paper, and lots of writing!

\paragraph{3. Pod Configuration Schema}

@@

\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily,caption={RR Debug Session Schema},captionpos=b]
apiVersion: v1
kind: Pod
metadata:
  name: rr-translation
  labels:
    purpose: translate-rr
spec:
  containers:
  - name: rr-test-container
    image: example-container-registry.com/sproj/...
    securityContext:
      capabilities:
        add:
        - SYS_PTRACE
  restartPolicy: OnFailure
\end{lstlisting}

@@

\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily,caption={Frontend Load Balancer Schema},captionpos=b]
apiVersion: v1
kind: Service
metadata:
  name: frontend-load-balancer
spec:
  selector:
    app: frontend
  type: LoadBalancer
  ports:
    - port: 3000
    targetPort: 3000
\end{lstlisting}

@@

\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RR Message Server},captionpos=b]
@sio.on('join_channel')
def join_channel(sid, data):
    sio.enter_room(sid, data['channel'])

@@

\paragraph{Building Example Pods}

@@

\begin{lstlisting}[language=Dockerfile,caption={RR Debug Session Hash Example---Dockerfile},captionpos=b]
FROM example-container-registry.com/sproj/...

@@

\begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily,caption={Example Startup Script},captionpos=b]
gcc -g -o hash hash.c
rr record ./hash
python3 rrtranslation.py $1
\end{lstlisting}

@@

\paragraph{The RR Translation Program}

@@

\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RR Translation Main},captionpos=b]
if __name__ == '__main__':
    channel = sys.argv[1]
    sio.connect('http://rr-message-server-load-balancer:8000')
    sio.emit('join_channel', {'channel': channel})
\end{lstlisting}

@@

\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RR Command Event Handler},captionpos=b]
@sio.on('rr_command')
def on_rr_command(data):

@@

\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily,caption={RRInterface},captionpos=b]
def get_full_rr_response(self, command):
    response = self.gdbmi.write(command)
    while(not(self.end(response) or self.exited(response))):
        response.extend(self.get_rr_response())
    return response

@@

We just need this in here for the ref for now. TODO.

>>>>

\paragraph{3. Pod Configuration Schema} \label{podschema}

@@

\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily\small,caption={RR Debug Session Schema},captionpos=b]
apiVersion: v1
kind: Pod
metadata:
  name: rr-translation
  labels:
    purpose: translate-rr
spec:
  containers:
  - name: rr-test-container
    image: example-container-registry.com/sproj/...
    securityContext:
      capabilities:
        add:
        - SYS_PTRACE
  restartPolicy: OnFailure
\end{lstlisting}

@@

\begin{lstlisting}[language=YAML,basicstyle=\linespread{0.5}\ttfamily\small,caption={Frontend Load Balancer Schema},captionpos=b]
apiVersion: v1
kind: Service
metadata:
  name: frontend-load-balancer
spec:
  selector:
    app: frontend
  type: LoadBalancer
  ports:
    - port: 3000
    targetPort: 3000
\end{lstlisting}

@@

\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={RR Message Server},captionpos=b]
@sio.on('join_channel')
def join_channel(sid, data):
    sio.enter_room(sid, data['channel'])

@@

\paragraph{Building Example Pods} \label{buildingchannel}

@@

\begin{lstlisting}[language=Dockerfile,basicstyle=\small,caption={RR Debug Session Hash Example---Dockerfile},captionpos=b]
FROM example-container-registry.com/sproj/...

@@

\begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily\small,caption={Example Startup Script},captionpos=b]
gcc -g -o hash hash.c
rr record ./hash
python3 rrtranslation.py $1
\end{lstlisting}

@@

\paragraph{The RR Translation Program} \label{joiningchannel}

@@

\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={RR Translation Main},captionpos=b]
if __name__ == '__main__':
    channel = sys.argv[1]
    sio.connect('http://rr-message-server-load-balancer:8000')
    sio.emit('join_channel', {'channel': channel})
\end{lstlisting}

@@

\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={RR Command Event Handler},captionpos=b]
@sio.on('rr_command')
def on_rr_command(data):

@@

\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={RRInterface},captionpos=b]
def get_full_rr_response(self, command):
    response = self.gdbmi.write(command)
    while(not(self.end(response) or self.exited(response))):
        response.extend(self.get_rr_response())
    return response

@@

The second major backend system consists of the Frontend/API Server
and database.  This system is responsible for serving the frontend
webapp, as well as processing all API requests, such as those to
authenticate users and create new RR Debug Sessions.  The builtin
React development server is used to serve the frontend webapp and
redirect API requests to the Flask API server during development.  For
both security and performance reasons this should be changed to a
combination of a more robust solution like Nginx \cite{nginx} and a
production suitable web server for production.  The relationships
between components and overall process will remain unchanged after
this migration.
\par

Since the entire webapp is one page, the process for serving it is
completely standard.  The server recieves a request for the index page
of the website, and returns the compiled React webapp that makes up
the homepage.  The process for API requests is more involved, and an
example is outlined below:

  \includegraphics[scale=1]{login_request}
  \centering
  \caption{API Request Process---Login}
  \label{rr:detailed}
\end{figure}

\begin{enumerate}
\item The client makes a POST request to the \lstinline{'/login'} URL.
  This request contains the necessary data to log in the user.
\item The frontend server forwards the request of a URL it does not
  recognize to the API server running in the same Pod on a different
  port.
\item The API server recognizes the \lstinline{'/login'} route, and
  communicates with the database to attempt to log in the user.
\item The API server returns the result of the login request, which is
  returned by the frontend server to the client.
\end{enumerate}

\subsubsection{Frontend/API Server Configuration}

The Frontend/API Server uses Yarn \cite{yarn} to manage packages.
When a Frontend/API Server Pod is deployed, the Pod's startup script
uses the \lstinline{yarn start} and \lstinline{yarn start-api}
commands to start the frontend server and API server respectively.
These scripts are defined in the server's \lstinline{package.json}
configuration file:\\

\begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily\small,caption={Frontend/API Server Configuration},captionpos=b]
  "start": "react-scripts start",
  "start-api": "cd api && flask run --no-debugger",
  .
  .
  .
  "proxy": "http://localhost:5000",
\end{lstlisting}

Also of note in \lstinline{package.json} is the instruction to proxy
the API server, which is running on port 5000.  This achieves the
automatic forwarding of API requests to the API server described
above.  Since API requests are proxied through the same address
serving the frontend webapp, there are no issues with cross-origin
requests.

\subsubsection{The API Server}

The API server is implimented using Flask.  The server program
consists of handlers for the various API routes and instances of two
classes which communicate with the database/Kubernetes API:
\lstinline{TranslationPodManager} and \lstinline{UserManager}.  The
full structure of the API Server program, \lstinline{api.py} is shown
below:

  \includegraphics[scale=1]{api_structure}
  \centering
  \caption{Structure of the API Server}
  \label{rr:detailed}
\end{figure}

Care was taken to abstract as much as possible in the design of the
API server and the classes it instantiates.  The two classes
instantiated by \lstinline{api.py}, \lstinline{TranslationPodManager}
and \lstinline{UserManager}, each instantiate another class which
interacts with the database, \lstinline{DatabaseController}, as an
instance variable.  \lstinline{TranslationPodManager} further
instantiates the Kubernetes Core V1 Python API to communicate with the
cluster in order to create/delete pods.  This process of abstraction
ensures that the database, underlying cluster structure, and
authentication methods can all change without significant changes
needing to be made to \lstinline{api.py}.
\par

The most significant improvement that could be made to this structure
would be to break the \lstinline{UserManager} and
\lstinline{TranslationPodManager} instances out into seperate programs
running on their own Pods in the cluster.  This would allow changes to
be made in those classes, say to update the process of deleting Pods,
without needing to update the entire Frontend/API Server Pod.  For the
time being, the relative simplicity of the API means that the added
complexity and overhead of implementing a method to communicate
between dedicated \lstinline{TranslationPodManager} and
\lstinline{UserManager} Pods and Pods running the Frontend/API Server
doesn't seem worthwhile.  This may change in the future.
\par

Currently supported API requests are as follows:

\begin{enumerate}
\item \lstinline{'/login'}: Logs in the user given by
  \lstinline{'name'}, or, (given the rather lax development security)
  creates a user if none matching \lstinline{'name'} exists.  Returns
  the name of the user.
\item \lstinline{'/channel'}: How users join RR Debug Sessions.  Binds
  the user given by \lstinline{'name'} to an RR Debug Session Pod
  matching \lstinline{'channel'} in the database, if such a Pod
  exists.  Returns the channel.
\item \lstinline{'/pods'}: How the client gets a list of RR Debug
  Session Pods the user is currently participating in.  Returns the
  channels for all RR Debug Session Pods who the user, given by
  \lstinline{'name'}, is currently bound to in the database.
\item \lstinline{'/examples}: How the client gets a list of example RR
  Debug Session Pod names.  Returns all names of example RR Debug
  Session Pods that exist in the database.
\item \lstinline{'/new'}: Creates a new RR Debug Session Pod based on
  the name given by \lstinline{'program'}.  Binds the user, given by
  \lstinline{'name'}, to the Pod.  Returns the channel of the Pod.
  This request is the most complex, and is covered in more detail
  below.
\item \lstinline{'/delete'}: Deletes a binding between the user given
  by \lstinline{'name'} and the RR Debug Session Pod given by
  \lstinline{'channel'}.  If there are no bindings left between the
  Pod and users (if all the users have left the debug session),
  deletes the Pod from the database and the cluster.  Returns
  \lstinline{True}.
\end{enumerate}

All the API request handlers listed above return an error message in
the case of an error occuring during the processing of an API request.
Since most errors that could occur would either be the result of
unanticipated issues on the part of the user, or of some unforseen
catostrophic internal error, informing users of error specifics would
be unhelpful.  Some \lstinline{TranslationPodManager} and
\lstinline{UserManager} functions throw specific errors in the event
of duplicate usernames, channels that do not exist, etc.  These errors
are caught and meaningful error messages are returned to the frontend
webapp, which can then pass them on to the user.
\par

To examine the process of processing an API request, it makes sense to
look at the most complex example, creating a new RR Debug Session Pod.
This example shows the process for processing an API POST request,
interacting with the database, and dynamically creating a Pod.  All
API requests follow a similar procedure.

\paragraph{Processing a POST Request---Extracting Information}

Below is the route handler for the \lstinline{'/new'} URL, as well as
the instantiation of \lstinline{TranslationPodManager} and
\lstinline{UserManager}.\\

\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={API Server New RR Debug Session Event Handler},captionpos=b]
tpm = podmanager.TranslationPodManager(
                 url='mongodb://database-load-balancer',
                 port=27017)
um = usermanager.UserManager(
                 url='mongodb://database-load-balancer',
                 port=27017)
app = Flask(__name__)

@app.route('/new', methods=['POST'])
def new():
    try:
        name = request.get_json()['name']
        program = request.get_json()['program']
        channel = tpm.create_pod([name], program)
        return {'channel': channel}
    except:
        return {'error': 'Internal Error'}
\end{lstlisting}

All API routes only support POST requests.  The route handler first
extracts the relavant information from the POST request body, in this
case \lstinline{name} and \lstinline{program}.  \lstinline{name}
always corresponds the user who is making the request's username, and
\lstinline{program} corresponds to the name of the example RR Debug
Session image to be used.
\par

\lstinline{new} then calls \lstinline{TranslationPodManager}'s
\lstinline{create_pod} function to create a new RR Debug Session Pod.
\lstinline{create_pod} can bind multiple users to a Pod when it is
created, so \lstinline{name} is passed inside a list.

\paragraph{The Database} Before drilling into \lstinline{create_pod},
it may be helpful to examine the MongoDB database which
\lstinline{TranslationPodManager} and \lstinline{UserManager} interact
with.  The database consists of three collections: \lstinline{users},
\lstinline{pods}, and \lstinline{examples}.  Below is an example
record from each collection:

  \includegraphics[scale=1]{database_example}
  \centering
  \caption{Database Record Examples}
  \label{rr:detailed}
\end{figure}

\lstinline{pods} and \lstinline{users} have a many-to-many
relationship.  Each user can be bound to an unlimited number of pods,
and vice versa.  The \lstinline{DatabaseController} class includes
functions to efficiently retrieve \lstinline{users} and
\lstinline{pods} given one another, and to create/destroy bindings
between \lstinline{users} and \lstinline{pods}.
\par

\lstinline{examples} are independent of \lstinline{users} and
\lstinline{pods}.  The continer registry URLs of RR Debug Session
example images are stored in the database so that new images can be
easily added to the collaborative debugger upon creation.
\par

MongoDB keeps \lstinline{_id}s unique.  In addition, the collaborative
debugger defines uniqueness constrains on \lstinline{users:name},
\lstinline{pods:channel}, and \lstinline{examples:name} when the
database is created.

% \begin{lstlisting}[basicstyle=\linespread{0.5}\ttfamily\small,caption={Frontend/API Server Configuration},captionpos=b]
% users:            pods:               examples:
%   _id ObjectId      _id ObjectId        _id ObjectId
%   name String       channel String      name String
%   pods Array        users Array         container String

% Example user:
% {
% 	"_id" : ObjectId("5fc14c622dc45448d67189b3"),
% 	"name" : "sammy",
% 	"pods" : [
% 		ObjectId("5fc14d5b2dc45448d67189b4"),
% 		ObjectId("5fc59acb2dc45448d67189b6")
% 	]
% }

% Example pod:
% {
% 	"_id" : ObjectId("5fc14d5b2dc45448d67189b4"),
% 	"channel" : "19452",
% 	"users" : [
% 		ObjectId("5fc14c622dc45448d67189b3"),
% 		ObjectId("5fc317d22dc45448d67189b5")
% 	]
% }

% Example example:
% {
% 	"_id" : ObjectId("5fc14d50b92a752776d3fd7d"),
% 	"name" : "hash",
% 	"container" : "example-container-registry.com/sproj/..."
% }
% \end{lstlisting}

\paragraph{Processing a POST Request---Dynamic Pod Creation}

To dynamically create a Pod, \lstinline{create_pod} gets the container
registry URL of the image to base the Pod on, creates the new Pod,
binds users to the Pod, and returns the channel the Pod was created
with.  The first part of the function is shown below:\\

\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={Pod Creation 1},captionpos=b]
def create_pod(self, names, program):
  examples = self.get_examples()
  image = None
  try:
    image = examples[program]
  except:
    raise NoSuchExampleError
\end{lstlisting}

\lstinline{create_pod} calls \lstinline{TranslationPodManager}'s
\lstinline{get_examples} method to get a list of RR Debug Session
example image names and container registry URLs from the database.  In
the event that the user has somehow pased a spurious image name or has
passed a name when there are no images, an exception is raised.
Otherwise, a RR Debug Session Pod schema is created using the image:\\

\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={Pod Creation 2},captionpos=b]
  dep = {'apiVersion': 'v1',
         'kind': 'Pod',
         'metadata': {'labels':
                      {'purpose': 'translate-rr'}},
         'spec': {'containers':
                  [{'image': image,
                    'name': 'rr-test-container',
                    'command': ['sh'],
                    'args': ['startup.sh'],
                    'securityContext':
                    {'capabilities':
                     {'add': ['SYS_PTRACE']}}}],
                    'restartPolicy': 'OnFailure'}}
\end{lstlisting}

This schema, represented as a dictionary in Python, is identical to
the YAML representation of the RR Debug Session Pod creation schema
shown in (\ref{podschema}) with the exception of \lstinline{args}
which will be updated further in a later step.  Finally,
\lstinline{create_pod} does the
work necessary to dynamically create a Pod:\\

\begin{lstlisting}[language=Python,basicstyle=\linespread{0.5}\ttfamily\small,caption={Pod Creation 3},captionpos=b]
  channel = self.dbc.add_pod(
            self.dbc.get_userids_by_name(names))
  dep['metadata']['name'] = channel
  dep['spec']['containers'][0]['args'].append(channel)
  resp = self.v1.create_namespaced_pod(
         body=dep, namespace='default')
  # Wait to return the channel until the pod is live
  # and read to recieve incomming communication
  status = self.v1.read_namespaced_pod_status(
           channel, 'default').status.container_statuses
  while status == None or status[0].state.running == None:
    time.sleep(1)
    status = self.v1.read_namespaced_pod_status(
             channel, 'default').status.container_statuses
  return channel
\end{lstlisting}

First, \lstinline{DatabaseController}'s \lstinline{add_pod} method is
called to insert a new \lstinline{pods} record into the database.
\lstinline{add_pod} randomly generates a new five digit
\lstinline{channel} for the Pod, using the uniqueness constraint
imposed on \lstinline{pods:channel} to ensure an unused
\lstinline{channel} is generated.  If no channels are open an error is
raised.  Otherwise, the users passed to \lstinline{add_pod} are bound
to the new Pod in the database, and \lstinline{channel} is returned.
This process has the effect of limiting the number of simultaneous RR
Debug Sessions that can be in use at the same time to 89999, and of
slowing as more channels are used.  Given the few current users, the
likelyhood of a collision when generating a new channel is low enough
that a more advanced method is unnecessary.
\par

Next, the name of the container is updated to \lstinline{channel}.  In
addition, the startup arguments are updated to include
\lstinline{channel}.  This is how \lstinline{channel} is passed to the
\lstinline{startup.sh} script and eventually used to join a channel on
the RR Message Server in \lstinline{rrtranslation.py}
(\ref{buildingchannel} \& \ref{joiningchannel}).
\par

The Kubernetes API is then finally used to create the Pod.  The Pod is
created in the default namespace.  This could be changed to logically
isolate dynamically and statically created Pods with minimal hassle,
since Kubernetes' DNS resolution can connect pods accros namespaces.
If Pod creation fails, an error will be thrown.  Otherwise, a loop is
used to wait until \lstinline{rrtranslation.py} is running inside the
Pod.  Once the Pod is ready, \lstinline{channel} is returned.

\paragraph{Processing a POST Request---Returning a Response}

Most request handlers, \lstinline{'/new'} included, simply return a
JSON serializable dicitionary containing whatever results the API
Server's instance of \lstinline{TranslationPodManager} or
\lstinline{UserManager} returned, or an error message.  The
\lstinline{'/new'} handler returns \lstinline{'channel': channel} so
that the frontend webapp's socket.io client can join the channel
corresponding to the RR Debug Session Pod that was just created.
